


%__________   DEDICATION  ______________________________
\vspace*{2in}
\begin{center}
	\textit{DEDICATION}
\end{center}

\begin{center}
	\textit{To my Father, Mois\'es Robles Alomar. To my Mother, Michele Pag\'an Phillips. To my Sister, Maritza Lee Robles Pag\'an. To my Partner, Patricia Michelle Miranda Col\'on.}
\end{center}
%____________________________________________________________

\newpage


%__________   ACKNOWLEDGMENTS  ______________________________
\vspace*{0.5in}
\begin{center}
\section*{ACKNOWLEDGMENTS}
\end{center}
%\addcontentsline{toc}{section}{ACKNOWLEDGMENTS}


\noindent First of all, I want to thank my family, friends, and peers who have been part of this journey. My parents, Mois\'es and Michele, my sister, Maritza, and my partner, Patricia, for the love and support they have given me during this process. Thankful to the THS team that helped make this project possible: Angelo, Fabiola, Celimar, Rosaidy, and Marianyelis. 

Lastly, I want to thank my advisor Manuel Rodríguez, for his mentorship and motivation during these last few years in college. You gave me the opportunity when no one else did and taught me more than I could have hoped. 

This research is supported by the United States (US) National Library of Medicine of the National Institutes of Health (NIH) under award number R15LM012275. The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH. 

%____________________________________________________________



